### 2022-10-05 manuscript/manuscript.tex
more intro

While background selection models do predict diversity levels

@@

even when sites under selection are sparse.

@@

Typically evolutionary genetic models consider

@@

>>>>

While background selection models do predict diversity levels  XXX

@@

@@

@@

Here, I step through a quick derivation of \textcite{Santiago1998-bs}.
Throughout, we assume random mating, hermaphroditic individuals, and a constant
population size.

The change in a neutral allele's frequency in one generation can be partitioned
into the three sources of stochasticity: the random associations with fitness
backgrounds (i.e. \emph{draft}), the non-heritable randomness in family size,
and the Mendelian noise from heterozygotes segregating. If we let $x_{0,i} \in
\{0, 1, 2\}$ be the number of neutral alleles individual $i$ in generation 0
carries, we can partition the neutral allele frequency change into the
underlying stochastic causes,

\begin{align}
  p_1 = \frac{1}{N} \sum_{i=1}^N \left( \frac{x_{0,i}}{2} k_i + \sum_{j=1}^{k_i} \delta_{0,i,j} \right)
\end{align}

where $k_i$ is the number of surviving gametes parent $i$ passes on, and
$\delta_{0,i,j}$ is a random term that encapsulates the noise due to Mendelian
segregation of heterozygotes. If parent $i$ is a homozygote ($x_{0,i} \in \{0,
2\}$), then $\delta_{0,i,j} = 0$, whereas if $x_{0,i} = 1$ then $\delta_{0,i,j}
= \pm \nicefrac{1}{2}$ with equal probability. This is because a heterozygous
parent will transmit half a neutral allele in expectation, but each round of
Mendelian segregation must pass on either 0 or 1 alleles, which the random $\pm
\nicefrac{1}{2}$ term imposes. There is the condition that $\nicefrac{1}{N}
\sum_i k_i = 2$, since each diploid parent must have two kids to maintain a
constant population size.

Then, noting that $p_0 = \nicefrac{1}{2N} \sum_{i=1}^N x_{0,i}$, the allele
frequency change is,

\begin{align}
  \Delta p_1 = p_1 - p_0 = \frac{1}{N} \sum_{i=1}^N \frac{x_{0,i}}{2} (k_i - 1) + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j}.
\end{align}

Next, we can partition the gamete number into heritable ($f_i$) and
non-heritable ($\varepsilon_i$) fitness components of each individual $i$, $k_i
= f_i + \varepsilon_i$. We define $f_i$ such that $\nicefrac{1}{N} \sum_i f_i =
2$ in a constant population, and $\varepsilon_i$ is noise term centered around
zero. Then, the neutral allele frequency change is,

\begin{align}
  \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
               \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
  + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
  % \Delta p_1 &= \frac{1}{2N} \sum_{i=1}^N x_{0,i} (f_i - 1) +
               % \frac{1}{2N} \sum_{i=1}^N x_{0,i} \varepsilon_i
  % + \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^{k_i} \delta_{0,i,j} \\
  \Delta p_1 &= \Delta_{_H} p_1 + \Delta_{_N} p_1 + \Delta_{_M} p_1
\end{align}

which we will derive the variance of through time (c.f. \cite{Santiago1995-hx}
equation 2 and \cite{Buffalo2019-qs} equation 30). This equation is very
general, and allows for any reproduction model with finite offspring variance
and can account for the impact of linked selection under a variety of selection
processes.

%%% OLD
